Co-operation between the sexes trumps selfish control of female meiotic drive. Or Sperm do not evolve to collaborate in female meiotic drive. Or Why does female meiosis allow the opportunity for exploitation by selfish sperm?

### 2013-12-10 writeup_and_notes/male_dependent_drive.tex
added eqn. to describe geno. freq evolution for simplest sperm dep. drive model

%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
\title{Sperm do not evolve to collaborate in  female meiotic drive}

@@

It is also not the case that sperm is mechanistically incapable of influencing the outcome of female meiosis.
Mechanistic evidence for this possibility comes from
\emph{C. elegans} the fact that premature deployment of the aster (a vital component of mitotic
machinery) provided by the sperm can disrupting MII meiotic segregation
in the egg, leading to a triploid zygote \citep{McNally2012}. 
Additionally, genetic evidence suggests that the transmission patterns
in heterozygous females many potentially depend on sperm haplotype. 
Specifically, the two best characterized female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
While such genetic evidence is suggestive, we note the difficulty in ruling out the alternative hypothesis of early selection    \citep[Give page numbers to discussion of this in][]{Burt2006}.

@@

>>>>

%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
\title{Co-operation between the sexes trumps selfish control of female
  meiotic drive. Or 
Sperm do not evolve to collaborate in female meiotic drive. Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? }

@@

It is also not the case that sperm is mechanistically incapable of influencing the outcome of female meiosis.
Mechanistic evidence for this possibility comes from
\emph{C. elegans} the fact that premature deployment of the aster (a vital component of mitotic
machinery) provided by the sperm can disrupting MII meiotic segregation
in the egg, leading to a triploid zygote \citep{McNally2012}. 
Additionally, genetic evidence suggests that the transmission patterns
in heterozygous females many potentially depend on sperm haplotype. 
Specifically, two female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
While such genetic evidence is suggestive, we note the difficulty in
ruling out the alternative hypothesis of early selection  \citep[pages
52-54][]{Burt2006}.

@@

\section*{Appendix Stuff}
$\Venus$
Genotype frequencies at birth $x_{AA},~x_{AB},~x_{BB}$ cost in
homozygotes 
\begin{equation}
x'_{BB} \propto w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
  x_{BB}^{\Mars} +\frac{1}{2} d
w_{AB} x_{AB}^{\Mars} )  
\end{equation}

\begin{equation}
x'_{AA} \propto w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
  w_{AB} x_{AB}^{\Mars}) 
\end{equation}

\begin{eqnarray}
x'_{AB} \propto  & w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
 w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
 x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) 
\end{eqnarray}
